Disk Winds Matter: Simulating the Optical Spectra of Cataclysmic Variables

### 2014-08-19 draft.tex
Version with some bullets, 19/08

\begin{document}
%% define bibstyle and other definitions
\bibliographystyle{aabib}
%% renew commands
\renewcommand{\labelitemi}{$-$}
\def\py{\textsc{Python} }
\def\tar{\textsc{Tardis} }
\def\cld{\textsc{Cloudy} }
\def\civ{C~\textsc{iv} }
\def\araa{ARAA}
\def\nat{Nature}
\def\apjl{ApJ Letters}
\def\aapr{AAPR}
\def\ssr{SSR}
\def\apj{ApJ}
\def\pasp{PASP}
\def\aap{A\&A}
\def\mnras{MNRAS}
\def\aj{AJ}
\def\rmxaa{RMXAA}

@@

The code is capable of operating in a number of different
regimes, both in terms of the scale of the system, and the 
shape and characteristic energies of the illuminating spectrum. It is 
first described by LK02, who initially developed a biconical CV 
wind model to synthesize UV spectra. 
Higginbottom et al.\nocite{higginbottom2013} (2013; hereafter H13) presented a 
benchmark model for broad absorption line (BAL) QSOs.
As part of this effort, the ionization scheme was improved so that 
the code was capable of treating the power-law X-ray spectrum
exhibited by AGN.
This involved using a spectral model the mean intensity in a cell
computed from MC estimators for, instead of using a dilute blackbody
approximation. Compton processes
were also incorporated into the heating and cooling balance.
In addition, \cite{simmacro2005} modeled
Brackett and Pfund line profiles in YSOs using \textsc{Python,} which included
implementing a `macro-atom' mode in order to correctly treat the more
complex transition matrix involved in recombination.

@@

Previously, \py used a two-level atom approximation \cite[see e.g.][]{mihalas}. This approximation is 
fast, and works well as a first approximation when treating so-called `resonance lines' (such as C\textsc{iv}) 
in which the excited electron is strongly coupled to the ground state.
However, in a number of situations, for example, when a level is primarily populated from above then,
this becomes a poor representation. 
To reproduce recombination lines and lines in which the lower level is not the ground state, 
it is therefore important to use an improved line transfer technique and level population solver.

By quantising matter into `macro-atoms', and radiant and kinetic 
energy into indivisible energy packets (r- and k- packets respectively), 
Lucy (2002, 2003\nocite{lucy2002, lucy2003}; hereafter L02, L03) showed that it is possible 
to asymptotically reproduce the emissivity of a gas in statistical equilibrium 
without simplifying the treatment of line transfer. 
It allows all possible transition paths from a given level and a full non-Local 
Thermodynamic Equilibrium (NLTE) solution for the level populations based on MC estimators. 
For a full description consult L02 and L03.

Macro-atoms have been used to model astrophysical plasmas in 
Wolf-Rayet star winds \cite{sim2004}, AGN disk winds \citep{simlong2008, tatum2012}
and Supernovae \citep{kasen2006, kerzendorfsim}.
Sim et al. (2005) used \py itself to apply this technique to
a Hydrogen-only model. We are now able to adopt a hybrid scheme. 
Any species in which have their level populations determined by
more complex interactions than resonance transitions
can be treated as full macro-atoms. The other species are treated as so-called `simple-ions',
where they still follow the indivisible packet constraint but use the two-level atom approximation.
This allows one to preserve the fast treatment of resonance lines while providing 
an improved representation of, for example, recombination, which is important when modeling features 
such as the Balmer lines and also recombination {\sl continuum} emission. 
Recombination is expected to be an important process in the optical
spectra of CVs, and they thus make a perfect application for Lucy's 
advances in MC methods.

@@

\subsubsection{Macro-atoms}
% In order to correctly model lines in which the lower level is not the ground state 
% (such as H$\alpha$ \& H$\beta$), and improve the treatment of recombination,
% we treat H \& He as macro-atoms. This is because 
% the levels responsible for the line emission are populated mostly from above.
% Therefore, 
we treat H \& He as macro-atoms. 
Macro-atom NLTE level populations and ion fractions are calculated by solving 
the statistical equilbrium equations between each pair of levels. The rate
between a lower level $l$ and upper level $u$ in an atom is given by

@@

\subsection{Atomic Data}
Atomic data is obtained from the same sources described in LK02 and updated in H13. 
The additional level and line information required to treat He as macro atom was 
obtained from \textsc{Topbase} \citep{topbase2005}. We treat H as a 20-level atom as in S05, where 
each level is defined by the principal quantum number n, which is appropriate due to
the degeneracy of each level in the H atom. He~\textsc{II} is treated in much the same way,
with 10 levels, but He~\textsc{I} has larger energy differentials between different l subshells
and triplet and singlet states. Thus, we still include levels up to $n=10$ but explicitly 
treat the $l$ and $s$ sub-orbitals as distinct levels instead of assuming they are perfectly `l-mixed',
which also allows the model to produce singlet and 
triplet lines, as seen in optical CV spectra.

@@

\begin{table}
\centering
\begin{tabular}{p{3cm}p{4cm}}
Model A \\
\hline Free Parameters 	&	 Value \\ 
\hline \hline 
$M_{WD}$ 	 &	 $0.8 M_{\odot}$ \\ 
$\dot{M}_{acc}$ 	 &	 $10^{-9}~M_{\odot}yr^{-1}$\\ 
$\dot{M}_{wind}$  &	$10^{-8}~M_{\odot}yr^{-1}$\\ 
$r_{min}$ 	&	 $4 R_{WD}$\\ 
$r_{max}$ 	&	 $12 R_{WD}$ \\ 
$\theta_{min}$ 	&	 $20.0^{\circ}$ \\ 
$\theta_{max}$ 	&	 $65.0^{\circ}$ \\ 
$\gamma$ 	&	 $1$ \\ 
$v_{\infty}$ 	&	 $3v_{esc}$ \\ 
$R_v$ 	        &	 $7\times10^{10}$cm \\ 
$\alpha$ 	&	 $1.5$ \\
\end{tabular}
\centering
\caption{
Wind geometry parameters used in the benchmark CV model.}
\label{wind_param}
\end{table}

@@

We hope that future simulations, coupled with a comparison to data, will 
actually address some of these problems, particularly since
the filtering effect of the wind may cause noteiceable effects
on the colour of the emergent spectrum \citep{hassall}. 
Here we use a blackbody illuminating spectrum
for the ionization cycles (see figure~\ref{sed}) and to compute our MC estimators.
During the spectral synthesis stage of the simulation we use stellar atmosphere models.
This is partly because they provide the most pessimistic prediction in terms of whether recombination emission in the 
wind can fill in the Balmer jump, due to the strong absorption beyond the Balmer edge.
For low temperatures, we use stellar atmospheres computed 
by Kurucz (reference), and for $T>50,000$~K we use models computed with 
\textsc{TLUSTY} \citep{tlusty}. The input spectra are created from these model
atmospheres using \textsc{Synspec}\footnote{http://nova.astro.umd.edu/Synspec43/synspec.html}.
%The temperature profile of our disk,
%with $\dot{M}_{acc} = 10^{-8}~M_{\odot}yr^{-1}$ can be seen in figure~\ref{t_profile}.

\begin{figure}
\centering
\includegraphics[width=0.45\textwidth]{figures/sed_figure.eps}
\caption{
The spectral energy distribution of the blackbody accretion
disk used in the ionization calculation (thick black line), 
with two boundary layer spectra also plotted.
The positions of the Hydrogen and Helium ionization edges 
are marked with dashed vertical lines.
%The effective temperature of the accretion disk
%in our model, as a function of radius.
}
\label{sed}
\end{figure}

\subsubsection{Boundary Layer}
The existence of a boundary layer in CVs arises naturally from the boundary condition
of the \cite{shakurasunyaev1973} model, since the material at the inner disk
edge must lose angular momentum in order to be deposited on
the surface of the compact object \cite[see e.g.][]{lyndenbell1974}.
Boundary layers are a possible source of
high energy photons which may be responsible for the photoionization of 
He~\textsc{ii}. 
\cite{pringlesavonije1979} predict temperatures of 
$200,000-500,000{\rm K}$ for the boundary layer,
but a number of studies have found difficulties
in producing the required ionization state of the wind
with such a hot, luminous boundary layer 
\citep[see e.g.][]{maucheraymond1987, drewverbunt1985}.
\cite{hoare1991}
use the Zanstra method to find boundary layer temperatures of
$\sim80,000{\rm K}$

@@

\begin{equation}
T_{bl} = \left[ \frac{L_{acc}}{8 \pi R_* H \sigma} ( 1 - \omega_*/\omega_k)^2 \right]^{1/4},
\label{bl}
\end{equation}

where $L_{acc}$ is the accretion luminosity, $R_*$ is the WD radius, $H$ is the boundary layer height
above the disk plane
and $\omega_*$ and $\omega_k$ are the angular velocities at the WD surface and inner disk edge respectively
\citep{fkrbook}. Note that we assume a cylindrical geometry such that the effective emitting
area is $4 \pi R_* H$. 
In the benchmark model, we follow LK02 in setting the boundary layer
luminosity to zero, and then investigate the effect
of boundary layers with temperatures of $80,000{\rm K}$ and $200,000{\rm K}$ 
on the ionization state of the wind and resulting
spectrum. See section 5 for a discussion.

@@

In fact, the wind has a Balmer jump in emission, but it is not prominent enough
to overcome the intrinsic absorption in the illuminating spectrum. 
Figure~\ref{cont} shows the total escaping spectrum, integrated
over all viewing angles, with the contribution from the disk and wind
also plotted. It shows that the lines are produced in the wind,
and also that the wind is producing a Balmer recombination
continuum, but one which is not yet sufficient to mask the absorption 
intrinsic to the disk SED. This offers a tantalising hint
that modest changes to the kinematics of the wind
may be sufficient to boost the continuum emission of the wind.
Furthermore, it confirms our suspicions that the improved
radiative transfer would cause enhanced recombination emission,
suggesting that, where possible, up to date treatments of
NLTE line transfer should be used in codes such as this.

The prominent P-Cygni profiles in e.g., C\textsc{iv} are still seen
in the UV spectra (see figure~\ref{spec}). In addition, the wind 
now contributes significant Lyman-$\alpha$ and
He~\textsc{ii}~$\lambda1640{\rm \AA}$  emission. 
We note that the ionization state of the wind 
is such that recombination to He~\textsc{ii} is significant, even
without a boundary layer source of higher energy photons.

@@

In addition to producing synthetic spectra, we are also able to 
analyse the physical properties of the outflow.
Figure~\ref{wind} shows various physical properties
of the wind geometry. The ionization structure of the wind
is very similar to the original LK02 model, as one might expect.
He is mostly ionized throughout the wind except for a small region near
the disk which is shielded from the photons produced by the hot inner disk.

C\textsc{iv} is the dominant Carbon ion throughout the wind, allowing
for an absorbing column which covers a large range of velocitieis, and is thus
sufficient to produce the required broad, deep and blue-shifted absorption lines
we expect to see in a large fraction of Nova-likes.

@@

where $n_i$ is the number density of the ion in question. 
The emission reaching an observer 
also depends on the optical depth along 
the line of sight, which can also significantly affect line profile shapes.
Therefore, any quantity which controls the density profile of the wind
will have a profound impact on both the 
amount of emission and the shape of the escaping line profiles.

As described in section 3, we can alter the density in the wind by 
adjusting the velocity law. We have therefore conducted a limited parameter 
sweep, exploring the effect of changing $R_v$, $\alpha$ and $\dot{M}_W$.

\subsection{The Balmer Lines}

\begin{figure} %fullpage
\mbox{
\subfigure{\includegraphics[width=0.5\textwidth]{figures/grid_alpha.eps}}
\quad
\subfigure{\includegraphics[width=0.5\textwidth]{figures/grid_beta.eps}}   
}
\caption{
% How the H$\alpha$ and H$\beta$ lines in our model change with the kinematic properties
% of the wind. The three separate panels show three different values of $\dot{M}_W$, 
% while $\alpha$ and $R_v$ are all gradually increased from top to bottom and left to right respectively.
% Larger values of these quantities increase the density at the base of the wind. There is a clear transition
% from double peaked to single peaked behaviour, and at the highest densities the wind
% essentially becomes neutral as the optically thick regions self-shield the wind from photoionizing radiation.
\bf{Each of these figures will highlight the next gen model in bold.
and contain a grid like the above showing a pair of spectral features in
the left and right panels. The grey shaded line will show the benchmark model
for easy comparison.  
The most promising parameters at the moment are acceleration exponent and launch radii.}
}
\label{halpha}
\end{figure} %fullpage

Figure~\ref{halpha} shows how the H$\alpha$ line changes with the kinematics 
of the wind at an inclination of $80^\circ$. For models with lower densities, the
line is double peaked, but we see a transition to a single peaked line when 
a high enough density is reach such that the line becomes optically thick.
This reproduces the behaviour predicted by \cite{MC96}, who suggested that
the poloidal velocity gradient in the wind can be larger than the gradient of 
the Keplerian velocity. This gradient can therefore modify the line profile shape
but requires the line to be optically thick.

@@

\label{balmerjump}
\begin{itemize}
\item that we are able to fill in the Balmer jump with the recombination continuum from the wind
if the EM is high enough
\item How this changes with inclination
\item Comparison to literature
\item Discussion of disk atmosphere treatment and its effect.
\end{itemize}

\begin{figure} %fullpage
\mbox{
\subfigure{\includegraphics[width=0.5\textwidth]{figures/grid_1640.eps}}
\quad
\subfigure{\includegraphics[width=0.5\textwidth]{figures/grid_4686.eps}}   
}
%%\includegraphics[width=0.5\textwidth]{figures/test.eps}
\caption{
The evolution of the He II lines with the kinematic properties
of the wind. 
% The panels are ordered in the same way as figure~\ref{halpha}. In general, increasing densities
% lead to a Balmer jump in emission, although as in the case of H$\alpha$ there is a transition
% when the wind becomes a neutral absorbing column.
}
\label{jump}
\end{figure}

\subsection{Helium Lines}
Discussion of Helium lines in the spectrum.
\begin{itemize}
\item He II 4686, He II 1640 strengths
\item Make the prediction that if 4686 is strong we might expect strong He II 3202. 
\item Discuss why it wouldn't have been seen yet (spectral coverage)
\item Helium I lines
\end{itemize}

\begin{figure} %fullpage
\mbox{
\subfigure{\includegraphics[width=0.5\textwidth]{figures/grid_balmer.eps}}
\quad
\subfigure{\includegraphics[width=0.5\textwidth]{figures/grid_uv.eps}}   
}
%%\includegraphics[width=0.5\textwidth]{figures/test.eps}
\caption{The evolution of the Balmer jump and C~\textsc{iv} line with the kinematic properties
of the wind. 
% The panels are ordered in the same way as figure~\ref{halpha}. In general, increasing densities
% lead to a Balmer jump in emission, although as in the case of H$\alpha$ there is a transition
% when the wind becomes a neutral absorbing column.
}
\label{jump}
\end{figure}

@@

\subsection{Boundary Layer}
\begin{itemize}
\item discuss the effect of 80000K and 200000K Boundary layers
\item Particularly referring to He II and CIV
\item Our findings support idea that 200000K BL would over ionize the wind- 
we favour 80,000K
\item I'd like to do a hypothetical Zanstra calculation here- state what
temperature our model would predict from the Zanstra method
\end{itemize}

>>>>

\begin{document}
%% define bibstyle and other definitions
\bibliographystyle{aabib}
%% renew commands
\renewcommand{\labelitemi}{$\bullet$}
\def\py{\textsc{Python} }
\def\tar{\textsc{Tardis} }
\def\cld{\textsc{Cloudy} }
\def\civ{C~\textsc{iv} }
\def\araa{ARAA}
\def\nat{Nature}
\def\apjl{ApJ Letters}
\def\aapr{AAPR}
\def\ssr{SSR}
\def\apj{ApJ}
\def\pasp{PASP}
\def\aap{A\&A}
\def\mnras{MNRAS}
\def\aj{AJ}
\def\rmxaa{RMXAA}

@@

The code is capable of operating in a number of different
regimes, both in terms of the scale of the system, and the 
shape and characteristic energies of the illuminating spectrum. It is 
first described by LK02, who initially developed a biconical CV 
wind model to synthesize UV spectra. 
Higginbottom et al.\nocite{higginbottom2013} (2013; hereafter H13) presented a 
benchmark model for broad absorption line (BAL) QSOs.
As part of this effort, the ionization scheme was improved so that 
the code was capable of treating the power-law X-ray spectrum
exhibited by AGN.
This involved using a spectral model the mean intensity in a cell
computed from MC estimators for, instead of using a dilute blackbody
approximation. Compton processes
were also incorporated into the heating and cooling balance.
In addition, \cite[][SDL05]{simmacro2005} modeled
Brackett and Pfund line profiles in YSOs using \textsc{Python,} which included
implementing a `macro-atom' mode in order to correctly treat the more
complex transition matrix involved in recombination.

@@

Previously, \py used a two-level atom approximation \cite[see e.g.][]{mihalas}. This approximation is 
fast, and works well as a first approximation when treating so-called `resonance lines' (such as C\textsc{iv}) 
in which the excited electron is strongly coupled to the ground state.
However, in a number of situations, for example, when a level is primarily populated from above, then
this becomes a poor representation. 
To reproduce recombination lines and lines in which the lower level is not the ground state, 
it is therefore important to use an improved line transfer technique and level population solver.

By quantising matter into `macro-atoms', and radiant and kinetic 
energy into indivisible energy packets (r- and k- packets respectively), 
Lucy (2002, 2003\nocite{lucy2002, lucy2003}; hereafter L02, L03) showed that it is possible 
to asymptotically reproduce the emissivity of a gas in statistical equilibrium 
without simplifying the treatment of line transfer. 
His macro-atom scheme allows all possible transition paths from a given level and a full non-Local 
Thermodynamic Equilibrium (NLTE) solution for the level populations based on MC estimators. 
For a full description consult L02 and L03.

Macro-atoms have been used to model astrophysical plasmas in 
Wolf-Rayet star winds \citep{sim2004}, AGN disk winds \citep{simlong2008, tatum2012}
and Supernovae \citep{kasen2006, kerzendorfsim}.
SDL05 used \py itself to apply this technique to
a Hydrogen-only model. We are now able to adopt a hybrid scheme. 
Any species in which have their level populations determined by
more complex interactions than resonance transitions
can be treated as full macro-atoms. The other species are treated as so-called `simple-ions',
where they still follow the indivisible packet constraint but use the two-level atom approximation.
This allows one to preserve the fast treatment of resonance lines while providing 
an improved representation of, for example, recombination, which is important when modeling features 
such as the Balmer lines and also recombination {\sl continuum} emission. 
Recombination is expected to be an important process in the optical
spectra of CVs, and they thus make a perfect application for Lucy's 
advances in MC methods.

@@

\subsubsection{Macro-atoms}
% In order to correctly model lines in which the lower level is not the ground state 
% (such as H$\alpha$ \& H$\beta$), and improve the treatment of recombination,
% we treat H \& He as macro-atoms. This is because 
% the levels responsible for the line emission are populated mostly from above.
% Therefore, 
we treat H \& He as macro-atoms. 
Macro-atom NLTE level populations and ion fractions are calculated by solving 
the statistical equilbrium equations between each pair of levels. The rate
between a lower level $l$ and upper level $u$ in an ion is given by

@@

\subsection{Atomic Data}
Atomic data is obtained from the same sources described in LK02 and updated in H13. 
The additional level and line information required to treat He as macro atom was 
obtained from \textsc{Topbase} \citep{topbase2005}. We treat H as a 20-level atom as in SDL05, where 
each level is defined by the principal quantum number n, which is appropriate due to
the degeneracy of each level in the H atom. He~\textsc{II} is treated in much the same way,
with 10 levels, but He~\textsc{I} has larger energy differentials between different l subshells
and triplet and singlet states. Thus, we still include levels up to $n=10$ but explicitly 
treat the $l$ and $s$ sub-orbitals as distinct levels instead of assuming they are perfectly `l-mixed',
which also allows the model to produce singlet and 
triplet lines, as seen in optical CV spectra.

@@

\begin{table}
\centering
\begin{tabular}{p{3cm}p{4cm}}
Model A \\
\hline Free Parameters 	&	 Value \\ 
\hline \hline 
$M_{WD}$ 	 &	 $0.8 M_{\odot}$ \\ 
$R_{WD}$ 	 &	 $7\times10^{8}$cm\\ 
$\dot{M}_{acc}$ 	 &	 $10^{-9}~M_{\odot}yr^{-1}$\\ 
$\dot{M}_{wind}$  &	$10^{-8}~M_{\odot}yr^{-1}$\\ 
$r_{min}$ 	&	 $4 R_{WD}$\\ 
$r_{max}$ 	&	 $12 R_{WD}$ \\ 
$\theta_{min}$ 	&	 $20.0^{\circ}$ \\ 
$\theta_{max}$ 	&	 $65.0^{\circ}$ \\ 
$\gamma$ 	&	 $1$ \\ 
$v_{\infty}$ 	&	 $3v_{esc}$ \\ 
%%$R_v$ 	        &	 $7\times10^{10}$cm \\ 
$R_v$ 	        &	 $100 R_{WD}$ \\ 
$\alpha$ 	&	 $1.5$ \\
\end{tabular}
\centering
\caption{
Wind geometry parameters used in the benchmark CV model.}
\label{wind_param}
\end{table}

@@

We hope that future simulations, coupled with a comparison to data, will 
actually address some of these problems, particularly since
the filtering effect of the wind may cause noticeable effects
on the colour of the emergent spectrum \citep{hassall}. 
Here we use a blackbody illuminating spectrum
for the ionization cycles (see figure~\ref{sed}) and to compute our MC estimators.
During the spectral synthesis stage of the simulation we use stellar atmosphere models.
This is partly because they provide the most pessimistic prediction in terms of whether recombination emission in the 
wind can fill in the Balmer jump, due to the strong absorption beyond the Balmer edge.
For low temperatures, we use stellar atmospheres computed 
by Kurucz (reference), and for $T>50,000$~K we use models computed with 
\textsc{TLUSTY} \citep{tlusty}. The input spectra are created from these model
atmospheres using \textsc{Synspec}\footnote{http://nova.astro.umd.edu/Synspec43/synspec.html}.
%The temperature profile of our disk,
%with $\dot{M}_{acc} = 10^{-8}~M_{\odot}yr^{-1}$ can be seen in figure~\ref{t_profile}.

\subsubsection{Boundary Layer}
The existence of a boundary layer in CVs arises naturally from the boundary condition
of the \cite{shakurasunyaev1973} model, since the material at the inner disk
edge must lose angular momentum in order to be deposited on
the surface of the compact object \cite[see e.g.][]{lyndenbell1974}.
Boundary layers are a possible source of
high energy photons which may be responsible for the photoionization of 
He~\textsc{ii}. 
\cite{pringlesavonije1979} predict temperatures of 
$200,000-500,000{\rm K}$ for the boundary layer,
but a number of studies have found difficulties
in producing the required ionization state of the wind
with such a hot, luminous boundary layer 
\citep[see e.g.][]{maucheraymond1987, drewverbunt1985}.
Hoare \& Drew (1991; HD91) \nocite{hoare1991}
use the Zanstra method to find boundary layer temperatures of
$\sim80,000{\rm K}$

@@

\begin{equation}
T_{bl} = \left[ \frac{L_{acc}}{8 \pi R_{WD} H \sigma} ( 1 - \omega_*/\omega_k)^2 \right]^{1/4},
\label{bl}
\end{equation}

where $L_{acc}$ is the accretion luminosity, $R_{WD}$ is the WD radius, $H$ is the boundary layer height
above the disk plane
and $\omega_*$ and $\omega_k$ are the angular velocities at the WD surface and inner disk edge respectively
\citep{fkrbook}. Note that we adopt a cylindrical geometry such that the effective emitting
area is $4 \pi R_* H$. 
In the benchmark model, we follow LK02 in setting the boundary layer
luminosity to zero, and then investigate the effect
of boundary layers with temperatures of $80,000{\rm K}$ and $200,000{\rm K}$ 
on the ionization state of the wind and resulting
spectrum. See section 5 for a discussion.

\begin{figure}
\centering
\includegraphics[width=0.45\textwidth]{figures/sed_figure.eps}
\caption{
The spectral energy distribution (SED) of the blackbody accretion
disk used in the ionization calculation (thick black line), 
with two boundary layer spectra also plotted.
The SED for an accretion disk
with the parameters used by HD91 in their Zanstra method calculations is shown in grey. 
It still assumes a Shakura-Sunyaev temperature profile, but uses
of $\dot{M}_{acc}= 5\times10^{-9}~M_{\odot}$yr$^{-1}$, and a slightly lower mass
white dwarf with $M_{WD}=0.7 M_{\odot}$ and $R_{WD} = 7\times10^{8}$cm.
The positions of the Hydrogen and Helium ionization edges 
are marked with dashed vertical lines.
%The effective temperature of the accretion disk
%in our model, as a function of radius.
}
\label{sed}
\end{figure}

@@

In fact, the wind has a Balmer jump in emission, but it is not prominent enough
to overcome the intrinsic absorption in the illuminating spectrum. 
Figure~\ref{cont} shows the total escaping spectrum, integrated
over all viewing angles, with the contribution from the disk and wind
also plotted. It shows that the lines are produced in the wind,
and also that the wind is producing a Balmer recombination
continuum, but one which is not yet sufficient to mask the absorption 
intrinsic to the disk SED. This offers a tantalising hint
that modest changes to the kinematics of the wind
may be sufficient to boost the continuum emission of the wind.
Furthermore, it confirms our suspicions that the improved
radiative transfer would cause enhanced recombination emission,
suggesting that, where possible, up to date treatments of
NLTE line transfer should be used in codes such as this.
The prominent P-Cygni profiles in e.g., C\textsc{iv} are still seen
in the UV spectra (see figure~\ref{spec}). In addition, the wind 
now contributes significant Lyman-$\alpha$ and
He~\textsc{ii}~$\lambda1640{\rm \AA}$  emission. 
% We note that the ionization state of the wind 
% is such that recombination to He~\textsc{ii} is significant, even
% without a boundary layer as a source of higher energy photons.

In addition to producing synthetic spectra, we are also able to 
analyse the physical properties of the outflow.
Figure~\ref{wind} shows various physical properties
of the wind geometry. The ionization structure of the wind
is very similar to the original LK02 model, as one might expect.
He is mostly ionized throughout the wind except for a small region near
the disk which is shielded from the photons produced by the hot inner disk.
C\textsc{iv} is the dominant Carbon ion throughout the wind, allowing
for an absorbing column which covers a large range of velocitieis, and is thus
sufficient to produce the required broad, deep and blue-shifted absorption lines
we expect to see in a large fraction of Nova-likes.

@@

@@

where $n_i$ is the number density of the ion in question.    
Therefore, any quantity which controls the density profile of the wind
will have a profound impact on both the 
amount of emission and the shape of the escaping line profiles.
The emission reaching an observer at a given frequency
also depends on the sobolev optical depth it experiences, which depends on both
the density and the velocity gradient projected
along the line of sight. This can also significantly affect line profile shapes.       
As described in section 3, we can alter the density in the wind by 
adjusting the velocity law (which also determines the velocity gradient).

We have therefore conducted a limited parameter 
sweep, exploring the effect of changing a few key parameters.  We focus on 
4 variables in order to explore the trends in our model. These are:

\begin{itemize}
 	\item  The acceleration length, $R_v$, which controls the characteristic scale
 	over which the wind accelerates.
 	\item The acceleration exponent, $\alpha$, which focuses the acceleration
 	around $R_v$. A higher value of $\alpha$ therefore leads to a slowly accelerating 
 	wind with a dense base.
 	\item The launch radii of the wind, $r_{min}$ \& $r_{max}$.
 	\item The temperature of the boundary layer, $T_{bl}$.
 \end{itemize}

We have found these to be important parameters in modifying 
the spectral features and ionization state of the wind.
In general, quantities which cause a denser wind base increase
the overal emission, although optical depth effects can clearly cause 
the relationship to become more complicated.

\subsection{Line Profiles: Shapes and Intensities}

\noindent\rule{16cm}{0.4pt}

{\bf
\noindent key points here:

\begin{itemize}
	\item producing single peaked line profiles is difficult to say the least
	\item with our high density models we can produce some single peaked line emission,
	mostly in Halpha and beta. the other lines seem to stay double peaked.
	\item this higher density causes (too much) collisionally excited line emission from CIV
	\item if the wind was more equatorial then this might help, but this isn't really supported by observations.
	\item we can definitely reproduce the behaviour predicted by Murray and Chiang
	\item Comment on the observed Balmer decrement being fairly flat, as is observed. Point out that
	this is necessary if H alpha is to be single peaked, as you need significant optical depths.
\end{itemize}
}

\noindent\rule{16cm}{0.4pt}

\begin{figure} %fullpage
\mbox{
\subfigure{\includegraphics[width=0.5\textwidth]{figures/grid_alpha_0.eps}}
\quad
\subfigure{\includegraphics[width=0.5\textwidth]{figures/grid_beta_0.eps}}   
}
\caption{
% How the H$\alpha$ and H$\beta$ lines in our model change with the kinematic properties
% of the wind. The three separate panels show three different values of $\dot{M}_W$, 
% while $\alpha$ and $R_v$ are all gradually increased from top to bottom and left to right respectively.
% Larger values of these quantities increase the density at the base of the wind. There is a clear transition
% from double peaked to single peaked behaviour, and at the highest densities the wind
% essentially becomes neutral as the optically thick regions self-shield the wind from photoionizing radiation.
\bf{Each of these figures will highlight the next gen model in bold.
and contain a grid like the above showing a pair of spectral features in
the left and right panels. The grey shaded line will show the benchmark model
for easy comparison.  
The most promising parameters at the moment are acceleration exponent and launch radii.}
}
\label{halpha}
\end{figure} %fullpage

Figure~\ref{halpha} shows how the H$\alpha$ and H$\beta$ lines change with the kinematics 
of the wind at an inclination of $80^\circ$. For H$\alpha$, in models with lower densities, the
line is double peaked, but we see a transition to a single peaked line when 
a high enough density is reach such that the line becomes optically thick.
This reproduces the behaviour predicted by \cite{MC96}, who suggested that
the poloidal velocity gradient in the wind can be larger than the gradient of 
the Keplerian velocity. This gradient can therefore modify the line profile shape.
It is clear from the behaviour of H$\beta$

In order to produce single-peaked line emission at high inclinations,
it is clear that a higher line opacity is needed at the red and blue wings relative to
the opacity at line centre. This can be achieved in a number of ways. A more equatorial 
wind would cause the line opacity at line centre to decrease
relative to the the wings of the line due to the increased radial component 
of the velocity gradient.

Our difficulty in producing single-peaked line emission from a full 
radiative transfer calculation has been encountered before. SDL05 found that a 
biconical disk geometry similar to that adopted here could not reproduce 
the observed single peaked profiles in YSOs.

@@

\label{balmerjump}

\noindent\rule{16cm}{0.4pt}

{\bf
\noindent key points here:

\begin{itemize}
	\item there is always recombination continuum emission from the wind, but it
	only because significant w.r.t disk when the wind accelerates more slowly.
	\item This higher density causes (too much) collisionally excited line emission from CIV
	\item supports findings from the line profile shapes conclusions above...suggests
	we need to alter the geometry in some respect, or find someway to suppress
	collisionally excited line emission from UV lines?
	\item perhaps one should increase the volume of the emitting area rather than the density,
	although this will also cause more thermal line emission...
	\item Bottom lines: filling in the Balmer jump can be done, and at the very minimum recombination
	emission from the wind can matter, especially if we believe that the lines come from the wind. 
	But to properly fill in the edge, one requires very high, possibly unrealistic, densities.
	\item make clear that the edge originally comes from the stellar atmosphere, and that
	the model we have used here is perhaps a little pessimistic.
	\item We predict that the Balmer jump becomes deeper relative to the continuum 
	\item This is probably due to foreshortening of the disk continuum but a fairly constant
	escaping Balmer continuum.
\end{itemize}
}

\noindent\rule{16cm}{0.4pt}

\begin{figure} %fullpage
\mbox{
\subfigure{\includegraphics[width=0.5\textwidth]{figures/grid_1640_0.eps}}
\quad
\subfigure{\includegraphics[width=0.5\textwidth]{figures/grid_4686_0.eps}}   
}
%%\includegraphics[width=0.5\textwidth]{figures/test.eps}
\caption{
The evolution of the He II lines with the kinematic properties
of the wind. 
% The panels are ordered in the same way as figure~\ref{halpha}. In general, increasing densities
% lead to a Balmer jump in emission, although as in the case of H$\alpha$ there is a transition
% when the wind becomes a neutral absorbing column.
}
\label{jump}
\end{figure}

\subsection{The Boundary Layer and Ionization State of the wind}

\noindent\rule{16cm}{0.4pt}

% important points from Hoare and Drew paper
% Zanstra method requires optically thick in He II Lyman continuum
% recombination lines are optically thin
% they note that He II n=2 edge opacity would cause more stringent limits. relevant to our edges.
% %

{\bf 
\noindent key points here:

\begin{itemize}
	\item we don't necessarily need a Boundary layer to ionize Helium II
	\item a Boundary layer of 200,000K overionizes Carbon IV
	\item we therefore favour lower boundary layer temperatures, as supported
	by the upper limits of Hoare and Drew 1991.
	\item we expect a boundary layer to matter more in lower accretion
	rate systems.
\end{itemize}
}

\noindent\rule{16cm}{0.4pt}

We find in all our models that He II line emission is significant, and that 
in most cases He  III is the dominant Helium ion in the wind, making recombination
to He II an important process. This occurs without the presence of a boundary 
layer. However, as a boundary layer is theoretically predicted 
it is still important to investigate the effects of introducing
one on the ionization state of the wind, and in particular
the resultant line emission from \heiiuv, \heiiuv and \civ.

We introduce boundary layers with temperatures of 80,000K and 200,000K,
commonly proposed values in the literature. We find that

It is also instructive here to conduct a hypothetical 
Zanstra method calculation, following the method of \cite{hoare1991}. 
We adopt their definitions of the Zanstra ratios in the Helium lines (equations 5 and 6 
in their paper),

\begin{equation}
G (\lambda228, 1800 \AA) = 2.4\times 10^{-3} \frac{f(\lambda1640)}{f_*(1800\AA)}
\end{equation}
and

\begin{equation}
G (\lambda228, 1800 \AA) = 1.3\times 10^{-3} \frac{f(\lambda4686)}{f_*(1800\AA)}.\\
\end{equation}

Here f(line) is the line flux in erg~cm$^{-2}$ s$^{-1}$.

% Discussion of Helium lines in the spectrum.
% \begin{itemize}
% \item He II 4686, He II 1640 strengths
% \item Make the prediction that if 4686 is strong we might expect strong He II 3202. 
% \item Discuss why it wouldn't have been seen yet (spectral coverage)
% \item Helium I lines
% \end{itemize}

\begin{figure} %fullpage
\mbox{
\subfigure{\includegraphics[width=0.5\textwidth]{figures/grid_balmer_0.eps}}
\quad
\subfigure{\includegraphics[width=0.5\textwidth]{figures/grid_uv_0.eps}}   
}
%%\includegraphics[width=0.5\textwidth]{figures/test.eps}
\caption{The evolution of the Balmer jump and C~\textsc{iv} line with the kinematic properties
of the wind. 
% The panels are ordered in the same way as figure~\ref{halpha}. In general, increasing densities
% lead to a Balmer jump in emission, although as in the case of H$\alpha$ there is a transition
% when the wind becomes a neutral absorbing column.
}
\label{jump}
\end{figure}

@@

% \subsection{Boundary Layer}
% \begin{itemize}
% \item discuss the effect of 80000K and 200000K Boundary layers
% \item Particularly referring to He II and CIV
% \item Our findings support idea that 200000K BL would over ionize the wind- 
% we favour 80,000K
% \item I'd like to do a hypothetical Zanstra calculation here- state what
% temperature our model would predict from the Zanstra method
% \end{itemize}
